---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# siseanalytics


siseanalytics is a package to characterize the classes obtained after a clustering. It makes it possible to characterize the classes in a univariate way, the variables are taken individually. But also multivariate. In addition, the pacakge allows to have the evaluation measures.

In this tutorial we will see the concrete functionality of the package

* Installation
* Evaluation metrics
    - Object creation and display
    - Performance indicators
    - Comparaison of clustering results

* Univariate characterization
    - Characterization of the partition
    - Characterization of clusters

* Multivariate characterization



## Installation

In your script, install the package using these command lines :

```{r}
#install.packages("devtools")
library(devtools)
#install_github("CCCelestine/siseanalytics")
library(siseanalytics)

```

Access to datasets :

```{r}
data(df_test) # data frame test 357 rows and 8 variables
data(pred_reel_3c) # predicted and real values 3 class 
data(pred_reel_2c) # predicted and real values 2 class 
data(pred_reel_2c_bis) # predicted and real values 2 class 

```

## Univariate characterization

### Characterization of the partition

In this part, we assess how each variable contributes to the constitution of the partition.  
First of all, we will see the case of **qualitative variables**.  
Barplot is a graph that shows categorical variables with rectangular bars with heights or lengths proportional to the values they represent.
We can see in this first graph how the numbers of our categorical variable (here the sex) are distributed among the clusters (here small / medium / large).
```{r}
barplotYX(df_test, "sexe", "val_pred")
```
In this second barplot, we can see how the numbers of clusters are distributed among the categories of the categorical variable.

```{r}
barplotXY(df_test, "sexe", "val_pred")
```

Since we are in the presence of two categorical variables (the cluster variable is one), we can create a contingency table between both. This table provides a first approach to the composition of clusters.
tab.quanti.line() and tab.quanti.col() functions are used to display the contingency tables in percentage. The calculation of the proportions is made either in rows or in columns.

```{r}
tab.quali.ligne(df_test$val_pred,df_test$sexe)
tab.quali.col(df_test$val_pred,df_test$sexe)
```

We can also perform a Chi-Square Independence Test as well as calculate Cramer's V, which is a measure derived from Chi-Square. These results allow us to know if the two variables are related. Regarding Chi-square, the p-value must be less than alpha (often 5%) to be able to affirm that the variables are dependent. As for Cramer's V, 0 means no liaison and 1 means perfect liaison.

```{r}
khi2(df_test$sexe,df_test$val_pred)
vcramer(df_test$sexe,df_test$val_pred)
```

Next, we will approach the case of **quantitative variables**.  

The boxplot summarizes indicators of the position of the studied variable (median, quartiles, minimum, maximum or deciles). This plot is mainly used to compare the same variable in several populations.
Here we will compare the size of people in our three clusters.

```{r}
boxplot(df_test, "val_pred", "taille")

```

In this table, we find the means by variable and by cluster, called the conditional means. The eta indicator indicates the proportion of variance explained by the clusters.

```{r}
df_test=as.data.frame(df_test)
data_quanti=df_test[,3:5]
tab.quanti(data_quanti,df_test$val_pred)
```

### Characterization of the clusters

We will now compare the clusters with each other.

The resCluster () function displays a summary of the different indicators for a cluster. First, there is the proportion of individuals who are part of this cluster among all individuals. Then, there are two tables : one for the quantitative variables and one for the qualitative variables.
The first compares the mean of a variable for a cluster and the overall mean, while the other compares the proportion of a variable in a cluster and the proportion in the overall population. Both tables have another indicator : the test value. It makes it possible to distinguish the important variables in the interpretation of clusters.

```{r}
data=as.data.frame(df_test[,-c(1,2)])
resCluster(data,df_test$val_pred,"grand")

```


Beyond the calculated value of the VT, we will focus on the disparities and concomitances between
clusters. This can be visualized on a radar diagram for quantitative variables.

```{r}
radar(data_quanti,df_test$val_pred)

```

## Evaluation metrics

### Creating the metrics object

We create 3 objects using the EvalMetrics function. This function takes as input dataframe containing the real values and the values predicted by a clustering.

```{r}
Obj2c <- EvalMetrics(pred_reel_2c$val_reel,pred_reel_2c$val_pred)
Obj3c <- EvalMetrics(pred_reel_3c$val_reel,pred_reel_3c$val_pred)
Obj2c_bis <- EvalMetrics(pred_reel_2c_bis$val_reel,pred_reel_2c_bis$val_pred)
```

ggMatconf function display the confusion matrix under a ggplot graph

```{r}
ggMatConf(Obj2c)
ggMatConf(Obj3c)
```

Thanks to the overload of the print method we can display the different attributes of our object.
```{r}
print(Obj2c)
print(Obj3c)

```
### Performance indicators

The package allows to calculate different indicators to evaluate the results of clustering.


Indicators | Details
--- | --- 
Erreur  | Overall model performance
Accuracy | Overall model performance (1- erreur)
Précision | How accurate positive predictions are
Sensibilite | Truly positive observation coverage
Specificity | Coverage of truly negative observations
Balanced Accuracy| Overall performance of the model, when classes are unbalanced
F1 score| Hybrid indicator used for unbalanced classes

### Comparaison of clustering results


Finally in this clustering evaluation part, the pacakge allows you to compare two clustering results stored in two different objects

```{r}
compareRes(Obj2c,Obj2c_bis)
```

The output will be a ggplot chart confronting the indicators
